---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  cache = FALSE
)
```

[![R-CMD-check](https://github.com/Martin-Umpierrez/exeval/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/Martin-Umpierrez/exeval/actions/workflows/R-CMD-check.yaml)



```{r,echo=FALSE, warning=FALSE}
library(knitr)
```
# exeval <img align="right" src = "man/figures/logo_new.png" width="135px">

# exeval: An R-package for External Evaluation of popPKPD Models. 

`exeval` package performs the *external evaluation process* of a popPKPD model. The external evaluation is done using an independent dataset from which the original popPKPD model was developed.  Currently, model fit is based on  [mapbayr](https://github.com/FelicienLL/mapbayr) package. 

External evaluation for a single model from can be done with:

* a population PKPD model ([mrgsolve](https://github.com/metrumresearchgroup/mrgsolve) format),
* a data set with concentrations (NONMEM format)

## Installation

You can install the development version of exeval from [GitHub](https://github.com/) with:
```{r, eval = F}
install.packages("devtools")
devtools::install_github("Martin-Umpierrez/exeval")
```



