---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  cache = FALSE
)
```

[![R-CMD-check](https://github.com/Martin-Umpierrez/exeval/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/Martin-Umpierrez/exeval/actions/workflows/R-CMD-check.yaml)



```{r,echo=FALSE, warning=FALSE}
library(knitr)
```
# <img align="right" src = "man/figures/exeval_logohex.png" width="135px">

# exeval: An R-package for External Evaluation of popPKPD Models. 

`exeval` provides a reproducible workflow for *external evaluation* of
population pharmacokinetic (popPK) and pharmacokinetic-pharmacodynamic
(PKPD) models using independent validation datasets. MAP estimation is
currently powered by [mapbayr](https://github.com/FelicienLL/mapbayr).


The package supports model-informed evaluation workflows based on:

- Maximum a posteriori (MAP) estimation via `mapbayr`
- posterior model updating
- a priori and Bayesian forecasting simulations
- predictive performance metrics
- model comparison across candidate models
- graphical diagnostics and fit screening
 

Models can be supplied as:

- `mrgsolve` model code ([mrgsolve](https://github.com/metrumresearchgroup/mrgsolve))
- compiled `mrgmod` objects
- built-in curated models included in `exeval`

Input datasets should follow standard pharmacometric event-table structure
(e.g. ID, TIME, DV, AMT, EVID, CMT, OCC), with helper functions available
for data preparation.


## Installation

You can install the development version of exeval from [GitHub](https://github.com/) with:
```{r, eval = F}
install.packages("devtools")
devtools::install_github("Martin-Umpierrez/exeval")
```

