---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/logo_new",
  out.width = "100%",
  cache = FALSE
)
```

[![R-CMD-check](https://github.com/Nicolas-Schmidt/preDose2/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/Nicolas-Schmidt/preDose2/actions/workflows/R-CMD-check.yaml)

```{r,echo=FALSE, warning=FALSE}
library(knitr)
```
# preDose <img align="right" src = "man/figures/predose_hex_only.png" width="135px">

# preDose: An R-package for Robust External Evaluation of popPKPD Models. 

preDose is a free and open source package that automatize the process of *external evaluation process* using an independent dataset from which the original popPKPD model was developed.

Currently, the user can use the package  choose to use the package based on the  [mapbayr](https://github.com/FelicienLL/mapbayr) package


You can perform an external evaluation for a single model from :

* a population PKPD model 
(coded in [mrgsolve](https://github.com/metrumresearchgroup/mrgsolve),
* a data set with concentrations (NONMEM format)

## Installation

You can install the development version of preDose from [GitHub](https://github.com/) with:
```{r, eval = F}
install.packages("devtools")
devtools::install_github("Martin-Umpierrez/preDose")
```

```{r, setup, include=FALSE}
library(mapbayr)
library(mrgsolve)
```

```{r message=FALSE, include=FALSE, results='hide'}
devtools::load_all()
```

## Example

This document presents a case of study to illustrates the external evaluation of two population pharmacokinetic model of Tacrolimus:

- **Han et al. (2011)**  
  *Prediction of the tacrolimus population pharmacokinetic parameters according to CYP3A5 genotype and clinical factors using NONMEM in adult kidney transplant recipients.*  
  European Journal of Clinical Pharmacology, 69(1), 53–63.

- **Zuo et al. (2013)**  
  *Effects of CYP3A4 and CYP3A5 polymorphisms on tacrolimus pharmacokinetics in Chinese adult renal transplant recipients: a population pharmacokinetic analysis.*
  Pharmacogenet Genomics. 2013 May;23(5):251-61. doi: 10.1097/FPC.0b013e32835fcbb6. PMID: 23459029.


The external evaluation was performed using data from an internal study conducted between 2022 and 2024 by Umpierrez et al.

```{r example}
library(preDose)
## basic example code
```

#### 1) Properly code you model 

##### 1.1) Code your model in mrgsolve format. 

Models can be defined in two different ways:

- **Inline model code**, which can be stored and accessed globally within the project.
- **External `.cpp` files**, which can be read and compiled directly by *mrgsolve*.

Additionally, the package provides several pre-coded models that can be accessed using the `list_models()` function, allowing users to readily explore and apply available population pharmacokinetic models.

```{r Model, echo=T, message=FALSE, results='hide'}
Han_etal_test<-
  '$PROB
# One Comparment Model with first order absorption- Ka is FIXED
$GLOBAL
#define CP (CENT/iV)
$CMT  @annotated
EV   : Extravascular compartment
CENT : Central compartment#two compt model with first order absorption

$PARAM @annotated 
CL  :  24.13 : Clearance for CYP3A5*3*3
V  :  716 : central volume
KA  : 4.5 : absorption rate constant
ETA1 : 0 : IIVCl (L/h)
ETA2 : 0 : IIVV (L)

$PARAM @annotated @covariate
POD    : 0   : COV POST OPERATIVE DAY
HCT      : 0  : COV HCH
WT      : 0  : COV WT
CYP3A5      : 0  : Polimorfismo CYP3A5
OCC     : -99  : Occasion, shall be passed by dataset imported

$ODE
dxdt_EV = -iKA*EV;
dxdt_CENT = iKA*EV  - iCL*CP;

$MAIN
##CYP3A5 effect on Cl##

double HM = 1.186 ;  ####Rapid Metabolizer ####
double IM = 1.13 ;  ####Intermediate Metabolizer ####
double PM = 1 ;  ####Poor Metabolizer####

if(CYP3A5==1) double CL_EFFECT = HM ;
if(CYP3A5==2) CL_EFFECT = IM ;
if(CYP3A5==3) CL_EFFECT = PM ;

double CL_HCT1 = 1.3458 ; ##Effect of HCT< 33
double CL_HCT2 = 1.124 ;  ##Effect of HCT >33

if(HCT< 33) double CL_HCT = CL_HCT1 ;
if(HCT >= 33) CL_HCT = CL_HCT2 ;

double CL_POD = - 0.00762 ;

double iCL =  CL *exp(ETA(1) + ETA1)* pow(POD, CL_POD) * CL_EFFECT * CL_HCT ;  
double iV =  V *exp(ETA(2) + ETA2) * exp (0.355*WT/59.025) ;  
double iKA =  KA ;    


$OMEGA @name IIV 
0.248 
0.237

$SIGMA  @name SIGMA @annotated
ADD : 0 : ADD residual error
PROP : 0.16 : Proportional residual error


$TABLE
double IPRED = CENT/iV;
double DV = IPRED * (1 + PROP) ;

$CAPTURE @annotated
CP : Plasma concentration (mass/volume)
iCL :  Clearance
iV : :Central Volume
iKA : KA: absorption rate constant
EVID : EVENT ID
DV : PREDICCION
OCC: OCCASION

               '

```

#### 2) Import your external data

Import the data set in NM-TRAN-formatted datasets
Ensure that the dataset structure aligns with the required format for proper processing.

```{r external_data}
data("tacrolimus_pk1_kidney", package = "preDose")  # Cargar dataset desde el paquete
head(tacrolimus_pk1_kidney)  # Ver primeras filas
```

#### 3) External model evaluation with `exeval_ppk()`

External model evaluation is performed using the `exeval_ppk()` function.  
This function automates the full evaluation workflow, including:

- Calculation of individual parameters
- Model updating based on individual estimates
- Simulation of predicted concentrations
- Comparison between observed and predicted data

##### Detailed workflow

A step-by-step description of the external evaluation procedure  
(i.e., individual parameter estimation, model updating, and simulation)  
is provided in the corresponding **vignette**.

Please refer to the vignette for a detailed and reproducible explanation of each step.

```{r externaleval, echo=TRUE, message=FALSE, warning=FALSE, cache=TRUE}
res.1 <- exeval_ppk(model = Han_etal_test,
                    drug_name = "Tacrolimus",
                    model_name = "Han_etal_2011",
                    tool = "mapbayr",
                    data = subset(tacrolimus_pk1_kidney, ID <11),
                    evaluation_type= "Progressive",
                    assessment = 'Bayesian_forecasting')  # Cargar dataset desde el paquete
print(res.1)

```

#### 4) Make Some Important Plots to compare metrics
Several plot types are available to visualize model performance using the `metrics_plot()` function.
##### 4.1) Bias BarPlot 
```{r plot1, echo=TRUE, message=FALSE}

plot1 = metrics_plot(res.1,
             type = "bias_barplot")

print(plot1)

```
##### 4.2) Bias boxplot 

```{r plot2, echo=TRUE, message=FALSE}

plot2 = metrics_plot(res.1,
             type = "bias_boxplot")

print(plot2)

```

##### 4.3) Relative Error Distribution by OCC 

```{r plot3, echo=TRUE, message=FALSE}

plot3 = metrics_plot(res.1,
             type = 'error_plot')

print(plot3)

```

#### 5) Import Models and assess the predicitve performance
##### 5.1) New Models and Estimations

```{r new_scenarios, echo=TRUE, warning=FALSE, message=FALSE, cache=TRUE}
source("inst/model_examples/ZuoX_etal_2013.R")

res.2 <- exeval_ppk(model = ZuoX_etalfull_noCYP3A4,
                    drug_name = "Tacrolimus",
                    model_name = "Zuo_etal",
                    tool = "mapbayr",
                    data = subset(tacrolimus_pk1_kidney, ID <11),
                    evaluation_type= "Progressive",
                    assessment = 'Bayesian_forecasting')  # Cargar dataset desde el paquete
print(res.2)
```

##### 5.2) Compare Models
###### 5.2.1) By Plotting: 
1. Use `combine_metrics()` to generate a summary of all evaluation metrics across the tested models.  
2. Visualize and compare model performance using the `plot_combined()` function.
```{r plot_comparisson, echo=TRUE, message=FALSE, warning=TRUE}

###### Generate a summary of metrics for all tested models
model_list <- list(list(model_name="Han_etal", metrics_list=res.1$metrics),
                   list(model_name="Zuo_etal", metrics_list=res.2$metrics))

#### Use combine_metrics() function with the summary created
combined_results<- combine_metrics(model_list)

#### Make the Plot! 
plot_comparrison <- plot_combined(combined_results,
                                  'bias_barplot')

print(plot_comparrison)

```


###### 5.2.2) Select models according to a specific evaluation metric and threshold using select_best_models() function
The `select_best_models()` function selects the best models from a dataframe of combined metrics based on a specified ranking metric.  
It requires a dataframe containing model evaluation metrics and the name of the metric to use for ranking.  
Optionally, you can specify a particular occasion to focus on and the number of top models to select.

```{r model_selection, echo=TRUE, message=FALSE, warning=TRUE}

Best_fit <- select_best_models(combined_results, metric = "rBIAS",
                               top_n = 1)

print(Best_fit)

```
